---
title: "lab_06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, include  = FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidytext)
library(forcats)
library(tidyverse)
```

```{r}
# read dataset
mtsample <- read.csv("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv")
mts <- as_tibble(mtsample)
```

<strong> <h4> Question 1: What specialties do we have? </h4> </strong>

```{r}
ms <- mtsample %>%
  count(medical_specialty)

ms %>%
    arrange(desc(n)) %>%
knitr::kable()
```

<strong> There are `r nrow(specialties)` medical specialties. </strong>

```{r}
ms %>%
  top_n(10) %>%
  ggplot(aes(x = n, y = fct_reorder(medical_specialty, n ))) +
    geom_col()
```

<strong> The distribution is not uniform. </strong>


<strong> <h4> Question 2 </h4> </strong>

```{r}
mtsample %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

<strong> There are lots of stop words here, non-specific to medical text. </strong>

<strong> <h4> Question 3 </h4> </strong>
```{r}
mtsample %>%
  unnest_tokens(word, transcription) %>%
  anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  #use regular expression to avoid numbers
  filter(!grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
````

<strong> Removing those stop words let us have a better idea about what the text is about </strong>

<strong> <h4> Question 4 </h4> </strong>
```{r}
mtsample %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```

```{r}
mts %>%
  unnest_ngrams(trigram, transcription, n=3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(trigram, n))) +
  geom_col()
```

<strong> Trigrams have a few more medical word group than bigrams. </strong>

<strong> <h4> Question 5 </h4> </strong>
```{r}
ptbigram <- mtsample %>%
  unnest_ngrams(bigram, transcription, n = 2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  select(word1, word2) %>%
  filter(word1 == "patient" | word2 == "patient")
```

<strong> Words appearing before patient: </strong>
```{r}
ptbigram %>%
  filter(word2 == "patient") %>%
  count(word1, sort = TRUE) %>%
  anti_join(stop_words, by = c("word1" = "word")) %>%
  top_n(10) %>%
knitr::kable()
```

<strong> Words appearing after patient: </strong>
```{r}
ptbigram %>%
  filter(word1 == "patient") %>%
  count(word2, sort = TRUE) %>%
  anti_join(stop_words, by = c("word2" = "word")) %>%
  top_n(10) %>%
knitr::kable()
```

<strong> <h4> Question 6</h4> </strong>

```{r}
mtsample %>%
  unnest_tokens(word, transcription) %>%
  group_by(medical_specialty) %>%
  count(word, sort = TRUE) %>%
  filter(!(word %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5, n) %>%
  arrange(medical_specialty, desc(n)) %>%
knitr::kable()
```