---
title: "assignment_04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HPC

### Problem 1: Make sure your code is nice

Original Function calculating total row sums
```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}
```

Rewrite fun1 function to make it faster 
```{r}
fun1alt <- function(mat) {
  # YOUR CODE HERE
  rowSums(mat, na.rm=TRUE)
}
```

Original Function calculating cumulative sum by row
```{r}
# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}
```

Rewrite fun2 function to make it faster 
```{r}
fun2alt <- function(mat) {
  # YOUR CODE HERE
  t(apply(mat, 1, cumsum))
}
```

Set up data example
```{r}
# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)
```

Test the first function
```{r}
# Test for the first
summary(microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat),check = "equivalent"
), unit = "relative")
```

Test the second function
```{r}
# Test for the second
summary(microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), check = "equivalent"
), unit = "relative")
```

### Problem 2: Make things run faster with parallel computing

The following function allows simulating PI. 
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using parLapply() to make it run faster.
```{r}
library(parallel)
# YOUR CODE HERE
cl <- parallel::makePSOCKcluster(2L)

# PREPARING THE CLUSTER
clusterSetRNGStream(cl, 1231) # Equivalent to `set.seed(1231)`

system.time({
  # YOUR CODE HERE
  clusterExport(cl,"sim_pi")
  ans <- unlist(parallel::parLapply(cl,1:4000, sim_pi, n = 10000))
  print(mean(ans))
})

# Stop the cluster
parallel::stopCluster(cl)
```

## SQL

### Setup a temporary database by running the following chunk

```{r}
library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

### Question 1

How many many movies is there available in each rating catagory.

```{sql, connection=con}
SELECT rating, COUNT(*) AS "movie_count"
FROM film
GROUP BY rating
```

### Question 2

What is the average replacement cost and rental rate for each rating category.
```{sql, connection=con}
SELECT rating, AVG(replacement_cost) AS "average of replacement cost", AVG(rental_rate) AS "average of rental rate"
FROM film
GROUP BY rating
```

### Question 3

Use table film_category together with film to find the how many films there are with each category ID

```{sql, connection=con}
SELECT category_id, COUNT(*) AS movie_count
FROM film
INNER JOIN film_category
ON film.film_id = film_category.film_id
GROUP BY category_id
```

### Question 4

Incorporate table category into the answer to the previous question to find the name of the most popular category.

```{sql, connection=con}
SELECT category.name, COUNT(*) AS movie_count
FROM film
INNER JOIN film_category
ON film.film_id = film_category.film_id
INNER JOIN category
ON film_category.category_id = category.category_id
GROUP BY category.name
ORDER BY movie_count DESC
```
Sports is the most popular category

### Clean up
Run the following chunk to disconnect from the connection.
```{r}
dbDisconnect(con)
```