---
title: "assignment_02"
output: html_document
---

```{r}
# set up libraries
library(data.table)
library(tidyverse)
```

<h2> Data Wrangling </h2>

```{r}
# read datasets
individual <- read.csv("chs_individual.csv")
regional <- read.csv("chs_regional.csv")
```

```{r}
# check the dimension
dim(individual)
dim(regional)
```

```{r}
# merge datasets 
CHS_data <-merge(
  # Data
  x     = individual,      
  y     = regional, 
  # List of variables to match
  by.x  = "townname",
  by.y  = "townname", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  ) 
```

<h4> Question 1 </h4>
```{r}
# check the number of rows and columns for merge data 
dim(CHS_data)
```

<strong> Since the individual data contains 1200 rows and 23 variables, the regional data contains 12 rows and 
27 variables, so the merge data should have 1200 rows and (23+27-1) vairables, which is 49 variables. Our merge data (CHS_data) exactly contains 1200 rows and 49 variables. </strong>

```{r}
# check the which variable has missing data 
names(which(colSums(is.na(CHS_data))>0))
```

```{r}
# In the case of missing values, impute data using the average within the variables “male” and “hispanic.” 
CHS_data <- data.table(CHS_data)
CHS_data[, bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = TRUE)), by = .(male, hispanic)]
CHS_data[, fev_imp := fcoalesce(fev, mean(fev, na.rm = TRUE)), by = .(male, hispanic)]
```

```{r}
# check whether the miss data in the bmi_imp is replaced by average 
CHS_data[is.na(bmi), table(bmi_imp)]
CHS_data[!is.na(bmi)] %>%
  group_by(male,hispanic) %>%
  summarize(bmi = mean(bmi))

# check whether the miss data in the fev_imp is replaced by average 
CHS_data[is.na(fev), table(fev_imp)]
CHS_data[!is.na(fev)] %>%
  group_by(male,hispanic) %>%
  summarize(fev = mean(fev))
```

<h4> Question 2 </h4>

```{r}
# Create a new categorical variable named “obesity_level” using the BMI measurement 
CHS_data[, obesity_level := fifelse(bmi_imp <14, "underweight",
                        fifelse(bmi_imp >=14 & bmi_imp <= 22, "normal",
                               fifelse(bmi_imp >22 & bmi_imp <= 24, "normal", "obese")))]
```

```{r}
# create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category
summary(CHS_data$bmi_imp)
table(CHS_data$obesity_level)
```


<h4> Question 3 </h4>

```{r}
# Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.”
CHS_data[, smoke_gas_exposure := fifelse(smoke == 1 & gasstove == 1, "expose both",
                        fifelse(smoke == 1 & gasstove != 1, "expose second hand smoke only",
                               fifelse(smoke != 1 & gasstove == 1, "expose gas stove only", "neither")))]
```

```{r}
# check the total number of observations per category
table(CHS_data$smoke_gas_exposure)
```

<h4> Question 4 </h4>

```{r}
# Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

# By town
summary_town <- CHS_data[, .(
    fev_avg      = mean(fev_imp, na.rm=TRUE),
    fev_sd       = sd(fev_imp, na.rm=TRUE),
    asthma_avg   = mean(asthma, na.rm=TRUE),
    asthma_sd    = sd(asthma, na.rm=TRUE)
    ),
    by = townname
    ][order(townname)] %>% head(n = 4)
summary_town
```

```{r}
# By sex
summary_sex <- CHS_data[, .(
    fev_avg      = mean(fev_imp, na.rm=TRUE),
    fev_sd       = sd(fev_imp, na.rm=TRUE),
    asthma_avg   = mean(asthma, na.rm=TRUE),
    asthma_sd    = sd(asthma, na.rm=TRUE)
    ),
    by = male
    ][order(male)] %>% head(n = 4)
summary_sex
```

```{r}
# By obesity level
summary_obesity_level <- CHS_data[, .(
    fev_avg      = mean(fev_imp, na.rm=TRUE),
    fev_sd       = sd(fev_imp, na.rm=TRUE),
    asthma_avg   = mean(asthma, na.rm=TRUE),
    asthma_sd    = sd(asthma, na.rm=TRUE)
    ),
    by = obesity_level
    ][order(obesity_level)] %>% head(n = 4)
summary_obesity_level
```

```{r}
# By smoke_gas_exposure
summary_smoke_gas_exposure <- CHS_data[, .(
    fev_avg      = mean(fev_imp, na.rm=TRUE),
    fev_sd       = sd(fev_imp, na.rm=TRUE),
    asthma_avg   = mean(asthma, na.rm=TRUE),
    asthma_sd    = sd(asthma, na.rm=TRUE)
    ),
    by = smoke_gas_exposure
    ][order(smoke_gas_exposure)] %>% head(n = 4)
summary_smoke_gas_exposure
```




<h2> Looking at the Data (EDA) </h2> 

<h4> Question 1 </h4>

```{r}
# Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”
library(ggplot2)

facet_plot <- CHS_data %>% 
  ggplot(mapping = aes(x = bmi_imp, y = fev_imp)) + 
  geom_point(mapping = aes(color=townname)) + 
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  facet_wrap(~ townname, nrow = 1) +
  scale_x_discrete(guide = guide_axis(n.dodge=3))+ 
  labs(x = "BMI", y = "FEV", title = "Scatterplots with regression lines of BMI vs FEV by “townname”")
facet_plot
```

<strong> BMI and FEV are positively correlation </strong>

```{r}
# Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default

# BMI category
hist1 <- CHS_data %>%
  ggplot(aes(fev_imp, fill = obesity_level)) +
  geom_histogram(binwidth = 5) +
  scale_fill_discrete(name = "Obesity Level") +
  labs(x = "FEV", y = "Count", title = "Stacked histogram of FEV by BMI category")
hist1

# smoke/gas exposure
hist2 <- CHS_data[!is.na(smoke_gas_exposure)] %>%
  ggplot(aes(fev_imp, fill = smoke_gas_exposure)) +
  geom_histogram(binwidth = 5) +
  scale_fill_discrete(name = "Smoke Gas Exposure") +
  labs(x = "FEV", y = "Count", title = "Stacked histogram of FEV by Smoke Gas Exposure")
hist2
```

<strong> Both graphs show the nearly normal distribution. </strong>

<h4> Question 3 </h4>

```{r}
# Barchart of BMI by smoke/gas exposure
barchart <- CHS_data[!is.na(smoke_gas_exposure)] %>%
  ggplot(mapping = aes(y = bmi_imp, x = smoke_gas_exposure)) + 
  geom_col(mapping = aes(colour = smoke_gas_exposure, fill = smoke_gas_exposure)) +
  scale_x_discrete(guide = guide_axis(n.dodge=3))+
  labs(x = "Smoke Gas Exposure", y = "BMI", title = "Barchart of BMI by smoke/gas exposure")
barchart
```

<strong> People expose gas stove only have the biggest range in BMI, whereas people exlose second
hand smoke only have the smallest range in BMI </strong> 

<h4> Question 4 </h4>

```{r}
# Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category

# by BMI
stat1 <- CHS_data %>%
ggplot() + 
    stat_summary(mapping = aes(x = obesity_level, y = fev_imp),
    fun.min = min,
    fun.max = max,
    fun = median,) + 
  labs(x = "Obesity Level", y = "FEV", title = "Statistical summary graphs of FEV by BMI category")
stat1

# by smoke/gas exposure category
stat2 <- CHS_data[!is.na(smoke_gas_exposure)] %>%
ggplot() + 
    stat_summary(mapping = aes(x = smoke_gas_exposure, y = fev_imp),
    fun.min = min,
    fun.max = max,
    fun = median,) + 
  labs(x = "Smoke Gas Exposure", y = "FEV", title = "Statistical summary graphs of FEV by Smoke Gas Exposure")
stat2
```

<strong> Obese children have higher FEV than other two categories. When it comes to the smoke gas exposure,
all categories have the almost same median, but the ranges for children expose second hand smoke only and expose both are smaller </strong>


<h4> Question 5 </h4>

```{r}
# A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
library(leaflet)
# Generating a color palette
pal <- colorFactor(c('blue','red'), domain=CHS_data$pm25_mass)

leafletmap <- leaflet(CHS_data) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
    label = ~paste0(pm25_mass, ' C'), color = ~ pal(pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 2
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=pal, values=CHS_data$pm25_mass,
          title='PM2.5 mass')

leafletmap
```

<strong> Cities near Los Angeles (especially near Pasadena) have much higher PM2.5 mass, whereas cities near 
Santa Barbara (near sea) contain less PM2.5 mass </strong>

<h4> Question 6 </h4>

```{r}
# Use geometric objects
geo <- CHS_data[!is.na(smoke_gas_exposure)] %>%
  ggplot() + 
  geom_smooth(method = 'gam', mapping = aes(x = pm25_mass, y = fev_imp))+
  labs(x = "PM2.5 mass", y = "FEV", title = "Graph between PM2.5 mass and FEV")
geo
```

<strong> We could see that with the increase of PM2.5 mass, FEV decreases. meaning that they have the 
negative correlation </strong> 

<p></p>
<p></p>
<p></p>

<h3> Conclusion </h3>

<strong> What is the association between BMI and FEV (forced expiratory volume) </strong>

BMI and FEV have a positive relationship. Higher BMI, higher FEV 

<strong> What is the association between smoke and gas exposure and FEV? </strong>

Smoke and gaas exposure forms the normal distribution with FEV. When FEV is around the median value,
the number of children under four categories is at most. In addition, there is no big difference of
the relationship between FEV and different categories in smoke and gas exposure

<strong> What is the association between PM2.5 exposure and FEV </strong>

They have the negative correlation. More PM2.5 exposure, less FEV 
